---
title: "Fall W271 Lab 2"
author: "John Kenney"
date: "October 9, 2017"
output: pdf_document
---


###House Keeping
```{r}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)

#Libraries required 
library(car)
library(dplyr)
library(Hmisc)#Used by author for 3D plotting
library(ggplot2)

library(effsize) #Used to calculate Cohen's D for T-Test
library(aod)    #Used for effect size of the logit model
library(mcprofile) #Used for confidence intervals
```

##Exploratory Data Analysis
We begin with an exploration of the data. Let's first look at our data at a high level.

```{r}
my.data = read.csv("C:\\Users\\jkenney\\Dropbox\\UCB\\Fall 2017\\W271\\Labs\\Lab 2\\lab2data.csv", header = TRUE)
glimpse(my.data)
```

We have 1,000 observations, each with 12 associated variables.

```{r}
#X Variable
describe(my.data$X)
min(my.data$X)
max(my.data$X)
```

It looks like this is the ID of the respondent, since there are 1000 unique numbers.

```{r}
#Gender Variable
describe(my.data$Gender)
```

There is a reasonably equal proportion of men and women in the sample.

```{r}
#Class Year Variable
describe(my.data$Class.Year)
```

Here we have five classes for graduation year that look to be coded in ten year intervals: 

1. 1972-1981
2. 1982-1991
3. 1992-2001
4. 2002-2011
5. 2012-Present

It seems the proportion of respondents is not independent of the graduation class.  The more recently graduated students make up a greater proportion of the sample than older graduates.  This distribution looks to be ordinal.

```{r}
ggplot(my.data, aes(x = Class.Year)) + geom_bar(aes(y=..count..), fill = "#1268AE", colour = "black")+ ggtitle("Graduation Decade") + theme(plot.title = element_text(lineheight = 1, face = "bold"))
```

```{r}
#Marital Status Variable
describe(my.data$Marital.Status)
```

We have four classes for the Marital Status varible:

1. Divorced
2. Married
3. Single
4. Widowed

```{r}
ggplot(my.data, aes(x = Marital.Status)) + geom_bar(aes(y=..count..), fill = "#1268AE", colour = "black")+ ggtitle("Graduation Decade") + theme(plot.title = element_text(lineheight = 1, face = "bold"))
```
```{r}
#Marital Status Variable
describe(my.data$Major)
```

We have forty-five distinct classes for the Major!  Wow.  Let's take a look at a bar chart

```{r}
my.data.major.count = as.data.frame(sort(table(my.data$Major), decreasing = TRUE)[1:20])
ggplot(my.data.major.count, aes(x = Var1, y = Freq)) + geom_bar(stat = "identity", fill = "#1268AE", colour = "black")+ ggtitle("Top 20 Majors") + theme(plot.title = element_text(lineheight = 1, face = "bold")) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Major") + ylab("Count")
```


```{r}
#Next Degree Variable
summary(my.data$Next.Degree)
my.data.next.degree = as.data.frame(sort(table(my.data$Next.Degree), decreasing = TRUE)[1:20])
ggplot(my.data.next.degree, aes(x = Var1, y = Freq)) + geom_bar(stat = "identity", fill = "#1268AE", colour = "black")+ ggtitle("Top 20 Next Degrees") + theme(plot.title = element_text(lineheight = 1, face = "bold")) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Degree") + ylab("Count")
```

Now on to Event Attendance...
```{r}
describe(my.data$AttendenceEvent)
```
More in attendance than not with 60% attending an event.

Finally, on to giving, our dependent variables.

$ FY12Giving      <dbl> 50, 0, 100, 0, 0, 0, 0, 5, 0, 0, 0, 0, 10, 0, 0, 0, 0, 20, 100, 30, 100, 0, 0, 50, 50, 0, 0, 0, 0, 50, 500...
$ FY13Giving      <dbl> 51, 0, 0, 0, 0, 0, 0, 10, 0, 75, 0, 0, 0, 0, 0, 5, 0, 50, 160, 0, 75, 0, 0, 100, 75, 0, 0, 0, 0, 50, 500, ...
$ FY14Giving      <dbl> 51, 0, 100, 0, 0, 0, 0, 25, 0, 0, 0, 0, 0, 0, 0, 0, 0, 50, 200, 0, 0, 0, 0, 156, 50, 0, 0, 0, 0, 50, 1000,...
$ FY15Giving      <dbl> 0, 0, 100, 0, 0, 0, 0, 25, 0, 0, 0, 0, 10, 0, 0, 0, 0, 50, 150, 50, 0, 0, 0, 157, 80, 0, 0, 0, 500, 50, 15...
$ FY16Giving
```{r}
describe(my.data$FY12Giving)
describe(my.data$FY13Giving)
describe(my.data$FY14Giving)
describe(my.data$FY15Giving)
describe(my.data$FY16Giving)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

