---
title: "Lab 02"
author: "Alex Yang, John Kenney, Ram Balasubramanian"
subtitle: "w271"
date: "Oct 12, 2017"
output: pdf_document
fontsize: 11pt
geometry: margin=0.75in
---

#Introduction:  
We have been hired by a Private University to identify who among their Alumni are most likely to contribute towards the University's foundation in future years. The university has provided us with data on past contributions from graduates - data includes some demographic information (like gender, marital status etc.), some university specific information (graduation year, major of studies), and some post-graduation information (Alumni event attendance, historical contributions).  
## We have taken two approaches to the problem (named Beta-Hat and Y-Hat):  
##Approach Beta-Hat:
We will treat the problem as a "explanation" problem ($\hat{\beta}$).  The goal here is to figure out if and how much certain aspects of a person and their association with the university determines how much they will contribute to the university's foundation.  We will develop a regression model that considers the 2016 contributions as a variable that depends on one or more of the other data elements that have been provided.  The regression coefficients can then be interpreted as a measure of how much each aspect of a person influences their contributions.  

##Approach Y-Hat:  
We will treat the problem as a "prediction" problem ($\hat{y}$ problem).  Given all the data we have about a person and their past contributions, can we predict how much they will contribute in the future.  We will develop a model that aims to predict the 2016 contribution amounts for each person.  To evaluate the efficacy of our models, we will split the data into a "training" set and a "test" set. We will use the training data to estimate parameters for our prediction model and evaluate our model's prediction accuracy using the test set.  




```{r}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)

#Libraries required 
library(car)
library(plyr)
library(dplyr)
library(Hmisc)#Used by author for 3D plotting
library(ggplot2)

library(effsize) #Used to calculate Cohen's D for T-Test
library(aod)    #Used for effect size of the logit model
library(mcprofile) #Used for confidence intervals

library(data.table)
```

### Read Data

```{r}
dt<-fread("lab2data.csv")
describe(dt)
```

We do not have any missing values in the data; and there do not seem to be an obvious "data cleaning" that needs to be conducted.  We will conduct an Exploratory Data Analysis next.  

# Exploratory Data Analysis:  

##Univariate Analysis:  
Let's examine each variable first starting with the "variable of interest" - 2016 contributions.

FY16Giving: Given that the vast majority of people did not give in 2016 and the skewness of the data (with a few large contributions) - let's also look at the distribution after a log-transformation (this is something we may want to consider for our modeling purposes)  



```{r, fig.height=3, fig.width=8}
par(mfrow=c(1,2))
a=count(dt$FY16Giving)
hist(dt$FY16Giving, breaks ='fd', xlab="2016 Contribution", main= "Distribution of 2016 Contributions")
hist(log10(dt$FY16Giving), breaks =10, xlab="Log(2016 Contributions)", main="Distribution of Log(2016 Contributions)")
```
Let's look at a distribution of just the "givers" (i.e. take out the zero contributions) to get a better picture.  

```{r, fig.height=3, fig.width=8}
par(mfrow=c(1,2))
hist(dt[FY16Giving>0]$FY16Giving, breaks=20, main="Dist of 2016  Contributions \n Givers Only", xlab= "2016 Contributions")
hist(log(dt[FY16Giving>0]$FY16Giving), breaks=20, main="Dist of Log(2016 Contributions) \n Givers only", xlab= "Log(2016 Contributions)")

```

Let's group the 2016 contributions by the categories that the university is interested in. 

```{r}
dt$FY16GivingCat<-cut(dt$FY16Giving,c(0,1,100,250,500,200000),right=FALSE)
summary(dt$FY16GivingCat)
```



Do it for other years as well
```{r}
dt$FY15GivingCat<-cut(dt$FY15Giving,c(0,1,100,250,500,200000),right=FALSE)
dt$FY14GivingCat<-cut(dt$FY14Giving,c(0,1,100,250,500,200000),right=FALSE)
dt$FY13GivingCat<-cut(dt$FY13Giving,c(0,1,100,250,500,200000),right=FALSE)
dt$FY12GivingCat<-cut(dt$FY12Giving,c(0,1,100,250,500,200000),right=FALSE)
```

And do a scatterplot matrix

```{r}
plot(jitter(as.numeric(FY16GivingCat),2)~jitter(as.numeric(FY15GivingCat),1.5),data=dt,pch=20)
```

So there is strong correlation between giving in one year and another
Distribution of log give (minus the 0's)
```{r}
ggplot(dt[FY16Giving>0],aes(x=log10(FY16Giving))) +
  geom_histogram(aes(y = ..density..), binwidth = 0.4, fill="#0072B2", colour="black") +
  ggtitle("Log(base 10) Give") + 
  theme(plot.title = element_text(lineheight=1, face="bold"))

```
The distribution of log wages is somewhat normal, with the majority less than $100, and a tail that stretches into 4 digits

Log is base 10 because meaning is more inuitive than a natural log

```{r}
ggplot(dt,aes(factor(AttendenceEvent),as.numeric(FY16GivingCat))) +
  geom_boxplot(aes(fill=factor(AttendenceEvent))) +
  ggtitle("Giving Category by Attendence at Event") +
  geom_jitter() +
  scale_x_discrete(name="Attendence",labels=c("No","Yes"))+
  scale_y_continuous(name="Giving Category",breaks=1:5,labels=c("[0,1)","[1,100)","[100,250)",
                                                                "[250,500)","[500,2e+05)")) +
  theme(plot.title = element_text(lineheight=1, face="bold")) 
```
So there does seem to be a relationship between giving in 2016 and attendence at the alumni event
Try giving Category by marital status


```{r}
ggplot(dt,aes(factor(Marital.Status),as.numeric(FY16GivingCat))) +
  geom_boxplot(aes(fill=factor(Marital.Status))) +
  ggtitle("Giving Category by Marital Status") +
  geom_jitter() +
  scale_x_discrete(name="Marital Status",labels=c("Divorced","Married","Single","Widowed"))+
  scale_y_continuous(name="Giving Category",breaks=1:5,labels=c("[0,1)","[1,100)","[100,250)",
                                                                "[250,500)","[500,2e+05)")) +
  theme(plot.title = element_text(lineheight=1, face="bold")) 
```


there are too few data points for divorced and widowed, but it does seem that married people donate more than single people

On the other hand, the distriubtion seems more even for widowed and divorced- there are just as many widowed alumni who give large amounts as there are who give little to nothing. This implies that giving might be related to age. Married people, after all, tend to be older than single people. so Marital status might just be capuring the effect of age or class Year

```{r}
ggplot(dt,aes(factor(Class.Year),as.numeric(FY16GivingCat))) +
  geom_boxplot(aes(fill=factor(Class.Year))) +
  ggtitle("Giving Category by Class Year") +
  geom_jitter() +
  scale_x_discrete(name="Class Year")+
  scale_y_continuous(name="Giving Category",breaks=1:5,labels=c("[0,1)","[1,100)","[100,250)",
                                                                "[250,500)","[500,2e+05)")) +
  theme(plot.title = element_text(lineheight=1, face="bold")) 
```
Here, it indeed looks like the older the alumnus, the more likely he or she is to give money.Unfortunately, the boxplot by category misses a lot of data, so we can try to instead do a boxplot of the log of 2016 giving, minus the people who donated nothing
```{r}
ggplot(dt[FY16Giving>0],aes(factor(Class.Year),log10(FY16Giving))) +
  geom_boxplot(aes(fill=factor(Class.Year))) +
  ggtitle("Log(base10) Giving by Class Year") +
  geom_jitter() +
  scale_x_discrete(name="Class Year")+
  scale_y_continuous(name="Log(base 10) Giving") +
  theme(plot.title = element_text(lineheight=1, face="bold")) 
```

So for those who do donate, there seems to a clear trend that onlder alumni donate more


Is there a relationship in terms of gender?

```{r}
ggplot(dt,aes(Gender,as.numeric(FY16GivingCat))) +
  geom_boxplot(aes(fill=Gender)) +
  ggtitle("Giving Category by Class Year") +
  geom_jitter() +
  scale_x_discrete(name="Gender")+
  scale_y_continuous(name="Giving Category",breaks=1:5,labels=c("[0,1)","[1,100)","[100,250)",
                                                                "[250,500)","[500,2e+05)")) +
  theme(plot.title = element_text(lineheight=1, face="bold")) 
```
It's a little hard to tell- since it does look like men donate a bit more to the university, but it's also possible that there are fewer female alumni (fewer women attended college in 1972)
```{r}
ggplot(dt[FY16Giving>0],aes(Gender,log10(FY16Giving))) +
  geom_boxplot(aes(fill=Gender)) +
  ggtitle("Log(base10) Giving by Gender") +
  geom_jitter() +
  scale_x_discrete(name="Gender")+
  scale_y_continuous(name="Log(base 10) Giving") +
  theme(plot.title = element_text(lineheight=1, face="bold")) 
```
Of those that donate, men donate more. But we can't be sure this is significant as the median man donated less than the 75th percentile woman.

Try class year split by male and female:


```{r}
ggplot(dt,aes(factor(Class.Year),as.numeric(FY16GivingCat))) +
  geom_boxplot(aes(fill=Gender)) +
  ggtitle("Giving Category by Class Year") +
  geom_jitter() +
  scale_x_discrete(name="Class Year")+
  scale_y_continuous(name="Giving Category",breaks=1:5,labels=c("[0,1)","[1,100)","[100,250)",
                                                                "[250,500)","[500,2e+05)")) +
  theme(plot.title = element_text(lineheight=1, face="bold")) 
```
So split by class year, gender does not seem to be a significant factor, and in fact older women donate more than older men

```{r}
ggplot(dt[FY16Giving>0],aes(factor(Class.Year),log10(FY16Giving))) +
  geom_boxplot(aes(fill=Gender)) +
  ggtitle("Log(base10) Giving by Class Year and Gender") +
  geom_jitter() +
  scale_x_discrete(name="Class Year")+
  scale_y_continuous(name="Log(base 10) Giving") +
  theme(plot.title = element_text(lineheight=1, face="bold")) 
```
SO limiting it to just alumni who donated in 2016, we do see that even split by age, men donated more than women, except for the oldest alumni