---
title: "Lab 02"
author: "Alex Yang, John Kenney, Ram Balasubramanian"
subtitle: "w271"
date: "Oct 12, 2017"
output: pdf_document
fontsize: 11pt
geometry: margin=0.75in
---


#SECTION - 3  STATISTICAL MODELING:
Section 3: Statistical Modeling. Start the section summarizing the key results - what variables, if
any, are the key predictors of the year 2016 contribution? What are the key techniques you have
experimented? What method did you use in your final model? How did you choose the final model?
What model performance criteria did you use to choose the final model? What statistical infernece did
you perform? Explain them. Comment on statistical significance vs. economic significance.




```{r}
#model for determining whether someone will contribute or not

#2016 giving status as function of 2015
model.fit1a = glm(Giver16~(Giver15), data=dt, family = binomial(link="logit"))
summary(model.fit1a)
Anova(model.fit1a, test="LR")

#2016 giving status as function of prior year giving status
model.fit1b = glm(Giver16~(Giver15+ Giver14+Giver13+Giver12), data=dt, family = binomial(link="logit"))
summary(model.fit1b)
Anova(model.fit1b, test="LR")

#2016 giving status as function of class year, marital status, attendance (not using prior year)
model.fit1c = glm(Giver16~factor(Class.Year)+factor(Marital.Status)+AttendenceEvent , data=dt, family = binomial(link="logit"))
summary(model.fit1c)
Anova(model.fit1c, test="LR")

#2016 giving status as function of giving status in 2015, class year, marital status, attendance 
model.fit1d = glm(Giver16~Giver15+factor(Class.Year)+factor(Marital.Status)+AttendenceEvent , data=dt, family = binomial(link="logit"))
summary(model.fit1d)
Anova(model.fit1d, test="LR")

#2016 giving status as function of ALL prior year giving statuses , class year, marital status, attendance 
model.fit1e = glm(Giver16~Giver15+ Giver14+Giver13+Giver12+factor(Class.Year)+factor(Marital.Status)+AttendenceEvent , data=dt, family = binomial(link="logit"))
summary(model.fit1e)
Anova(model.fit1e, test="LR")

```


```{r}
#models for determining category - using all data 2016giving category as function of 
#2015 giving category 
library(nnet)
model.fit2a = multinom(FY16GivingCat~FY15GivingCat, data=dt)
summary(model.fit2a)
Anova(model.fit2a, test="LR")


#2015 giving category 
library(nnet)
model.fit2b = multinom(FY16GivingCat~FY15GivingCat+FY14GivingCat+FY13GivingCat+FY12GivingCat, data=dt)
summary(model.fit2b)
Anova(model.fit2b, test="LR")


```
